---
title: "D10 Dashboard Project - Public Health Priorities"
author: "D10 Data Solutions"
date: "09/09/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r, include=FALSE}
library(tidyverse)
library(here)
```
```{r, include=FALSE}
sdmd_combined_plus_zones <- 
  read_csv(here("clean_data/sdmd_combined_plus_zones.csv"))

sdmd_combined_plus_zones %>% 
distinct(age)
```




# Introduction | Jonathan

## Structure of presentation

## The Brief

## The Team





# Key findings - Life expectancy | Derek

## The data
- Data set used
- Source
- Why it was chosen

## Trend
- Overview
- Plot
- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
- Plot
- Findings / key areas for business to focus on



# Key findings - Smoking | John Paul

## The data
- Data set used
- Source
- Why it was chosen

## Trend
- Overview
- Plot
- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
- Plot
- Findings / key areas for business to focus on



# Key findings - Drug abuse | Jonathan

## The data
- Scottish Drug Misuse 
- Source
- Why it was chosen

## Trend by gender  

```{r, message=FALSE}
sdmd_combined_plus_zones %>% 
  filter(sex != "All") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>%
  filter(year != "2013/14") %>% 
  arrange(year) %>% 
  group_by(year, sex) %>% 
  summarise(number_assessed_total = sum(number_assessed)) %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = sex, colour = sex)) +
  geom_point() +
  geom_line()
```
  
## Trend by age group  

```{r}
sdmd_combined_plus_zones %>% 
  filter(age != "All" &
         age != "Unknown" &
         year != "2013/14") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>% 
  arrange(year) %>% 
  group_by(year, age) %>% 
  summarise(number_assessed_total = sum(number_assessed), .groups = "drop") %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = age, colour = age)) +
  geom_point() +
  geom_line()
```


## Rank
- Overview
```{r, message=FALSE}

```

- Findings / key areas for business to focus on / disparity between areas


## Conclusion | Jonathan
- Recap brief
- Recap of the main focus points for each of the three areas

# Dashboard Demonstration | Prathiba & Paddy


# Questions