---
title: "D10 Dashboard Project - Public Health Priorities"
author: "D10 Data Solutions"
date: "09/09/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r, include=FALSE}
library(tidyverse)
library(here)
```
```{r, include=FALSE}
sdmd_combined_plus_zones <- 
  read_csv(here("clean_data/sdmd_combined_plus_zones.csv"))

life_expectancy_clean <- 
  read_csv(here("clean_data/life_expectancy.csv"))

sdmd_combined_plus_zones %>% 
distinct(age)
```




# Introduction | Jonathan

## Structure of presentation

## The Brief

## The Team





# Key findings - Life expectancy | Derek

## The Data

- What is life expectancy and why was it chosen?

- The source of the data

- Why was this data selected?

## Trend - D&G

- Life expectancy over time for Dumfries & Galloway
```{r, message=FALSE}
life_expectancy_clean %>%
  filter(type == "Local Authority",
         name == "Dumfries and Galloway",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code %in% c("1991-1993", "1993-1995", "1995-1997", "1997-1999", 
                   "1999-2001", "2001-2003", "2003-2005", "2005-2007", 
                   "2007-2009", "2009-2011", "2011-2013", "2013-2015",
                   "2015-2017", "2017-2019")) %>%
  ggplot() +
  aes(x = date_code, y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(
    x = "\nYear",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy for Dumfries & Galloway")
```
## Trend - Scotland

- Life expectancy over time for Scotland
```{r,message=FALSE}
life_expectancy_clean %>%
  filter(type == "Scotland",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code %in% c("1991-1993", "1993-1995", "1995-1997", "1997-1999", 
                   "1999-2001", "2001-2003", "2003-2005", "2005-2007", 
                   "2007-2009", "2009-2011", "2011-2013", "2013-2015",
                   "2015-2017", "2017-2019")) %>%
  ggplot() +
  aes(x = date_code, y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(
    x = "\nYear",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy for Scotland")
```

## Rank - Local Authority

- Life expectancy by local authority in the period 2017-2019
```{r, message=FALSE}
life_expectancy_clean %>% 
  filter(type == "Local Authority",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code == "2017-2019") %>%
  ggplot() +
  aes(x = reorder(name, le_value), y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    x = "\nLocal Authority",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy by Local Authority")
```
## Key Findings

- In Scotland in 2017-2019, the life expectancy at birth was 77.1 years for males, and 81.1 years for females.
- Life expectancy since 1991-1993 has increased but has largely remained unchanged since 2013-2015.
- Female life expectancy is lowest at birth in Glasgow (78.5 years) and highest in East Renfrewshire (84 years).
- Male life expectancy is lowest at birth in Glasgow (73.6 years) and highest in East Dunbartonshire (80.5 years).
- The majority of Scotland's council areas have seen a slow down in life expectancy since 2013-2015, with some seeing a decrease.

# Key findings - Smoking | John Paul

## The data
- The data set used for this was:
Smoking - Scottish Survey Core Questions

- The data was sourced from:
https://statistics.gov.scot/resource?uri=http%3A%2F%2Fstatistics.gov.scot%2Fdata%2Fsmoking-sscq

- Why it was chosen
Originally, Mental Wellbeing was selected for the project but the data quantity
was not enough for our purposes.  We chose Smoking as two of the Scottish 
Government's National Performance Framework (NPF) National Indicators are 
relevant to smoking. There is a strong Public Health focus on reducing the 
proportion of adults who smoke.

## Trend
- Overview
The overview is generally optimistic.  Whilst numbers are still higher than
anyone would like, the long term trend shows a steady reduction in all age groups, 
especially the youngest age group of 16-35.

- Plot

- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
There are huge variations around the country, whether they be measured by 
NHS Health Board or the more granular Local Authority areas.  Poverty is one 
factor along with an urban/rural divide.  Essentially the areas with the lowest
proportion of smokers are more rural, and more affluent.  The opposite is also true.

- Plot

- Findings / key areas for business to focus on
More work needs to be in less affluent areas, especially in the traditional 
industrial areas,  to further reduce the rate of smoking.


# Key findings - Drug abuse | Jonathan

## The data
- Scottish Drug Misuse 
- Source
- Why it was chosen

## Trend by gender  

```{r, message=FALSE}
sdmd_combined_plus_zones %>% 
  filter(sex != "All") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>%
  filter(year != "2013/14") %>% 
  arrange(year) %>% 
  group_by(year, sex) %>% 
  summarise(number_assessed_total = sum(number_assessed)) %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = sex, colour = sex)) +
  geom_point() +
  geom_line()
```
  
## Trend by age group  

```{r}
sdmd_combined_plus_zones %>% 
  filter(age != "All" &
         age != "Unknown" &
         year != "2013/14") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>% 
  arrange(year) %>% 
  group_by(year, age) %>% 
  summarise(number_assessed_total = sum(number_assessed), .groups = "drop") %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = age, colour = age)) +
  geom_point() +
  geom_line()
```


## Rank
- Overview
```{r, message=FALSE}

```

- Findings / key areas for business to focus on / disparity between areas


## Conclusion | Jonathan
- Recap brief
- Recap of the main focus points for each of the three areas

# Dashboard Demonstration | Prathiba & Paddy


# Questions