---
title: "D10 Dashboard Project - Public Health Priorities"
author: "D10 Data Solutions"
date: "09/09/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r, include=FALSE}
library(tidyverse)
library(here)
```
```{r, include=FALSE}
sdmd_combined_plus_zones <- 
  read_csv(here("clean_data/sdmd_combined_plus_zones.csv")) 

smoking <- read_csv(here("clean_data/smoking.csv"))

life_expectancy_clean <- 
  read_csv(here("clean_data/life_expectancy.csv"))

sdmd_combined_plus_zones %>% 
distinct(age)
```




# Introduction | Jonathan

## Structure of presentation

## The Brief

## The Team





# Key findings - Life expectancy | Derek

## The Data

- What is life expectancy and why was it chosen?

- The source of the data

- Why was this data selected?

## Trend - D&G

- Life expectancy over time for Dumfries & Galloway
```{r, message=FALSE}
life_expectancy_clean %>%
  filter(type == "Local Authority",
         name == "Dumfries and Galloway",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code %in% c("1991-1993", "1993-1995", "1995-1997", "1997-1999", 
                   "1999-2001", "2001-2003", "2003-2005", "2005-2007", 
                   "2007-2009", "2009-2011", "2011-2013", "2013-2015",
                   "2015-2017", "2017-2019")) %>%
  ggplot() +
  aes(x = date_code, y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(
    x = "\nYear",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy for Dumfries & Galloway")
```
## Trend - Scotland

- Life expectancy over time for Scotland
```{r,message=FALSE}
life_expectancy_clean %>%
  filter(type == "Scotland",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code %in% c("1991-1993", "1993-1995", "1995-1997", "1997-1999", 
                   "1999-2001", "2001-2003", "2003-2005", "2005-2007", 
                   "2007-2009", "2009-2011", "2011-2013", "2013-2015",
                   "2015-2017", "2017-2019")) %>%
  ggplot() +
  aes(x = date_code, y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(
    x = "\nYear",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy for Scotland")
```

## Rank - Local Authority

- Life expectancy by local authority in the period 2017-2019
```{r, message=FALSE}
life_expectancy_clean %>% 
  filter(type == "Local Authority",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code == "2017-2019") %>% 
  ggplot() +
  aes(x = reorder(name, le_value), y = le_value, group = sex, colour = sex) +
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(
    x = "\nLocal Authority",
    y = "Life Expectancy (years)\n",
    fill = "Sex"
  ) +
  ggtitle("Life Expectancy by Local Authority")
```
## Key Findings

- In Scotland in 2019-2019, the life expectancy at birth was 77.1 years for males, and 81.1 years for females.
- Life expectancy since 199-1993 has increased but has largely remained unchanged since 2013-2015.
- Female life expectancy is lowest at birth in Glasgow (78.5 years) and highest in East Renfrewshire (84 years).
- Male life expectancy is lowest at birth in Glasgow (73.6 years) and highest in East Dunbartonshire (80.5 years).
- The majority of Scotland's council areas have seen a slow down in life expectancy since 2013-2015, with some seeing a decrease.


# Key findings - Smoking | John Paul

## The data
- The data set used for this was:
Smoking - Scottish Survey Core Questions

- The data was sourced from:
statistics.gov.uk

- Why it was chosen
We chose Smoking as two of the Scottish Government's National Performance 
Framework (NPF)  National Indicators are relevant to smoking. 

There is a strong Public Health focus on reducing the proportion of adults who smoke.

## Trend
- Overview

The overview is optimistic.  

Whilst numbers are still higher than anyone would like, the long term trend 
shows a steady reduction in all age groups.

- Plot

```{r}
smoking %>%
  filter(smokes == "Yes",
         sex == "All") %>%
  
  arrange(date_code) %>% 
  group_by(date_code, age) %>% 
  summarise(avg_rate = mean(sm_percent)) %>% 
  

  ggplot(aes(x = date_code, y = avg_rate, group = age, colour = age)) +
  geom_point() +
  geom_line() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   labs(
    x = "\nYear",
    y = "Current Smokers (%)\n"
)
```

## Rank
- Overview

Where you live could affect whether you smoke or not.  

In some areas you are 300% more likely to smoke than in other areas.

- Plot

```{r}
#  Highest Smoking rates per Local Authority area, in a 1 year period
smoking %>% 
  filter(type == "Local Authority",
         type_of_tenure == "All",
         age == "All",
         household_type == "All",
         sex == "All",
         long_term_condition == "All",
         smokes == "Yes",
         date_code == "2019"
  ) %>%
 group_by(name, date_code, sm_percent) %>% 
  select(name, smokes, sm_percent) %>% 
  arrange(desc(sm_percent)) %>% 
head() %>% 
   ggplot(aes(x = reorder(name, sm_percent), y = sm_percent)) +
  geom_col() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   labs(
    x = "\nLocal Authority",
    y = "2019 Smoking Rate (%)\n")
```


```{r}
#  Lowest Smoking rates per Local Authority area, in a 1 year period
smoking %>% 
  filter(type == "Local Authority",
         type_of_tenure == "All",
         age == "All",
         household_type == "All",
         sex == "All",
         long_term_condition == "All",
         smokes == "Yes",
         date_code == "2019"
  ) %>%
 group_by(name, date_code, sm_percent) %>% 
  select(name, smokes, sm_percent) %>% 
  arrange(sm_percent) %>% 
  head() %>% 
   ggplot(aes(x = reorder(name, sm_percent), y = sm_percent)) +
  geom_col() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   labs(
    x = "\nLocal Authority",
    y = "2019 Smoking Rate (%)\n")
```

- Findings / key areas for business to focus on / disparity between areas

Where you live dictates how likely you are to smoke.

## Map
- Overview

There are huge variations around the country, whether they be measured by 
NHS Health Board or the more granular Local Authority areas.   

Essentially the areas with the lowest proportion of smokers are more affluent. The 
opposite is also true. 
 

- Findings

More work needs to be in less affluent areas to further reduce the rate of smoking.


# Key findings - Drug abuse | Jonathan

## The data
- Scottish Drug Misuse 
- Source
- Why it was chosen

## Trend by gender  

```{r, message=FALSE}
sdmd_combined_plus_zones %>% 
  filter(sex != "All") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>%
  filter(year != "2013/14") %>% 
  arrange(year) %>% 
  group_by(year, sex) %>% 
  summarise(number_assessed_total = sum(number_assessed)) %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = sex, colour = sex)) +
  geom_point() +
  geom_line()
```
  
## Trend by age group  

```{r}
sdmd_combined_plus_zones %>% 
  filter(age != "All" &
         age != "Unknown" &
         year != "2013/14") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>% 
  arrange(year) %>% 
  group_by(year, age) %>% 
  summarise(number_assessed_total = sum(number_assessed), .groups = "drop") %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = age, colour = age)) +
  geom_point() +
  geom_line()
```


## Rank
- Overview
```{r, message=FALSE}

```

- Findings / key areas for business to focus on / disparity between areas


## Conclusion | Jonathan
- Recap brief
- Recap of the main focus points for each of the three areas

# Dashboard Demonstration | Prathiba & Paddy


# Questions