---
title: "D10 Dashboard Project - Public Health Priorities"
author: "D10 Data Solutions"
date: "09/09/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r, include=FALSE}
library(tidyverse)
library(here)
```
```{r, include=FALSE}
sdmd_combined_plus_zones <- 
  read_csv(here("clean_data/sdmd_combined_plus_zones.csv")) 

smoking <- read_csv(here("clean_data/smoking.csv"))

sdmd_combined_plus_zones %>% 
distinct(age)
```




# Introduction | Jonathan

## Structure of presentation

## The Brief

## The Team





# Key findings - Life expectancy | Derek

## The data
- Data set used
- Source
- Why it was chosen

## Trend
- Overview
- Plot
- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
- Plot
- Findings / key areas for business to focus on



# Key findings - Smoking | John Paul

## The data
- The data set used for this was:
Smoking - Scottish Survey Core Questions

- The data was sourced from:
statistics.gov.uk

- Why it was chosen
We chose Smoking as two of the Scottish Government's National Performance 
Framework (NPF)  National Indicators are relevant to smoking. 

There is a strong Public Health focus on reducing the proportion of adults who smoke.

## Trend
- Overview

The overview is optimistic.  

Whilst numbers are still higher than anyone would like, the long term trend 
shows a steady reduction in all age groups.

- Plot

```{r}
smoking %>%
  filter(smokes == "Yes",
         sex == "All") %>%
  
  arrange(date_code) %>% 
  group_by(date_code, age) %>% 
  summarise(avg_rate = mean(sm_percent)) %>% 
  

  ggplot(aes(x = date_code, y = avg_rate, group = age, colour = age)) +
  geom_point() +
  geom_line() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   labs(
    x = "\nYear",
    y = "Current Smokers (%)\n"
)
```

## Rank
- Overview

Where you live could affect whether you smoke or not.  

In some areas you are 300% more likely to smoke than in other areas.

- Plot

```{r}
#  Highest Smoking rates per Local Authority area, in a 1 year period
smoking %>% 
  filter(type == "Local Authority",
         type_of_tenure == "All",
         age == "All",
         household_type == "All",
         sex == "All",
         long_term_condition == "All",
         smokes == "Yes",
         date_code == "2019"
  ) %>%
 group_by(name, date_code, sm_percent) %>% 
  select(name, smokes, sm_percent) %>% 
  arrange(desc(sm_percent)) %>% 
head() %>% 
   ggplot(aes(x = reorder(name, sm_percent), y = sm_percent)) +
  geom_col() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   labs(
    x = "\nLocal Authority",
    y = "2019 Smoking Rate (%)\n")
```


```{r}
#  Lowest Smoking rates per Local Authority area, in a 1 year period
smoking %>% 
  filter(type == "Local Authority",
         type_of_tenure == "All",
         age == "All",
         household_type == "All",
         sex == "All",
         long_term_condition == "All",
         smokes == "Yes",
         date_code == "2019"
  ) %>%
 group_by(name, date_code, sm_percent) %>% 
  select(name, smokes, sm_percent) %>% 
  arrange(sm_percent) %>% 
  head() %>% 
   ggplot(aes(x = reorder(name, sm_percent), y = sm_percent)) +
  geom_col() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   labs(
    x = "\nLocal Authority",
    y = "2019 Smoking Rate (%)\n")
```

- Findings / key areas for business to focus on / disparity between areas

Where you live dictates how likely you are to smoke.

## Map
- Overview

There are huge variations around the country, whether they be measured by 
NHS Health Board or the more granular Local Authority areas.   

Essentially the areas with the lowest proportion of smokers are more affluent. The 
opposite is also true. 
 

- Findings

More work needs to be in less affluent areas to further reduce the rate of smoking.


# Key findings - Drug abuse | Jonathan

## The data
- Scottish Drug Misuse 
- Source
- Why it was chosen

## Trend by gender  

```{r, message=FALSE}
sdmd_combined_plus_zones %>% 
  filter(sex != "All") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>%
  filter(year != "2013/14") %>% 
  arrange(year) %>% 
  group_by(year, sex) %>% 
  summarise(number_assessed_total = sum(number_assessed)) %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = sex, colour = sex)) +
  geom_point() +
  geom_line()
```
  
## Trend by age group  

```{r}
sdmd_combined_plus_zones %>% 
  filter(age != "All" &
         age != "Unknown" &
         year != "2013/14") %>% 
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>% 
  arrange(year) %>% 
  group_by(year, age) %>% 
  summarise(number_assessed_total = sum(number_assessed), .groups = "drop") %>% 
  ggplot(aes(x = year, y = number_assessed_total, group = age, colour = age)) +
  geom_point() +
  geom_line()
```


## Rank
- Overview
```{r, message=FALSE}

```

- Findings / key areas for business to focus on / disparity between areas


## Conclusion | Jonathan
- Recap brief
- Recap of the main focus points for each of the three areas

# Dashboard Demonstration | Prathiba & Paddy


# Questions