---
title: "D10 Dashboard Project - Public Health Priorities"
author: "D10 Data Solutions"
date: "09/09/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

# Title Page



# Introduction | Jonathan

## The Brief

## The Team

## Structure of presentation



# Key findings - Life expectancy | Derek

## The data
- Data set used
- Source
- Why it was chosen

## Trend
- Overview
- Plot
- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
- Plot
- Findings / key areas for business to focus on



# Key findings - Smoking | John Paul

## The data
- Data set used
- Source
- Why it was chosen

## Trend
- Overview
- Plot
- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
- Plot
- Findings / key areas for business to focus on



# Key findings - Drug abuse | Jonathan

## The data
- Data set used
- Source
- Why it was chosen

## Trend
- Overview
- Plot
- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
- Plot
- Findings / key areas for business to focus on



# Dashboard Demonstration | Prathiba & Paddy



## Conclusion | TBC
- Recap brief
- Recap of the main focus points for each of the three areas

# Questions