---
title: "D10 Dashboard Project - Public Health Priorities for Scotland"
author: "D10 Data Solutions"
date: "09/09/2021"
output:
  ioslides_presentation:
        widescreen: true
        css: styles.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r, include=FALSE}
library(tidyverse)
library(here)
```
```{r, include=FALSE}
sdmd_combined_plus_zones <-
  read_csv(here("clean_data/sdmd_combined_plus_zones.csv"))

life_expectancy_clean <-
  read_csv(here("clean_data/life_expectancy.csv"))

color_theme <- function() {
  theme(

    plot.background = element_rect(fill ="white"),
    plot.title = element_text(size = rel(2)),
    plot.title.position = "plot",

    panel.border = element_rect(colour = "blue", fill = NA, linetype = 1),
    panel.background = element_rect(fill = "white"),
    panel.grid =  element_line(colour = "grey85", linetype = 1, size = 0.5),

    axis.text = element_text(colour = "blue", face = "italic", size = 14),
    axis.title.y = element_text(colour = "#1B732B" ,  size = 14, angle = 90),
    axis.title.x = element_text(colour = "#1B732B" ,  size = 14),


    legend.box.background = element_rect(),
    legend.box.margin = margin(6, 6, 6, 6)

  )
}

```




# Introduction | Jonathan

## The Team

- Derek - Life Expectancy SME
- John Paul - Smoking SME
- JP - Drug Abuse SME
- Prathiba - R Shiny Dashboard Architect
- Paddy - Spatial Data Guru





# Key findings - Life expectancy | Derek

## The Data

- What is life expectancy and why it was chosen?

Life expectancy is a statistical measure of the average time an organism is expected to live, based on the year of its birth, its current age and any other demographic factors including sex and geographic area.

- The source of the data

statistics.gov.uk

- Why was this data selected?

## Trend - Dumfries & Galloway

```{r, message=FALSE, fig.width=9,fig.height=5}
life_expectancy_clean %>%
  filter(type == "Local Authority",
         name == "Dumfries and Galloway",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code %in% c("1991-1993", "1993-1995", "1995-1997", "1997-1999",
                   "1999-2001", "2001-2003", "2003-2005", "2005-2007",
                   "2007-2009", "2009-2011", "2011-2013", "2013-2015",
                   "2015-2017", "2017-2019")) %>%
  ggplot() +
  aes(x = date_code, y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(
    x = "\nYear",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy for Dumfries & Galloway") +
  color_theme()
```

## Trend - Scotland

```{r,message=FALSE, fig.width=9,fig.height=5}
life_expectancy_clean %>%
  filter(type == "Scotland",
         simd_quintiles == "All",
         urban_rural_classification == "All",
         sex != "All",
         age == 0,
         date_code %in% c("1991-1993", "1993-1995", "1995-1997", "1997-1999",
                   "1999-2001", "2001-2003", "2003-2005", "2005-2007",
                   "2007-2009", "2009-2011", "2011-2013", "2013-2015",
                   "2015-2017", "2017-2019")) %>%
  ggplot() +
  aes(x = date_code, y = le_value, group = sex, colour = sex) +
  geom_line() +
  geom_ribbon(aes(ymax = le_upper_ci, ymin = le_lower_ci), alpha = 0.25, colour = NA) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.5)) +
  labs(
    x = "\nYear",
    y = "Life Expectancy (years)\n",
    colour = "Sex"
  ) +
  ggtitle("Life Expectancy for Scotland") +
  color_theme()
```

## Rank
- Overview
- [plot placeholder]
- Findings / key areas for business to focus on / disparity between areas


# Key findings - Smoking | John Paul

## The data
- The data set used for this was:
Smoking - Scottish Survey Core Questions

- The data was sourced from:
https://statistics.gov.scot/resource?uri=http%3A%2F%2Fstatistics.gov.scot%2Fdata%2Fsmoking-sscq

- Why it was chosen
Originally, Mental Wellbeing was selected for the project but the data quantity
was not enough for our purposes.  We chose Smoking as two of the Scottish
Government's National Performance Framework (NPF) National Indicators are
relevant to smoking. There is a strong Public Health focus on reducing the
proportion of adults who smoke.

## Trend
- Overview
The overview is generally optimistic.  Whilst numbers are still higher than
anyone would like, the long term trend shows a steady reduction in all age groups,
especially the youngest age group of 16-35.

- [plot placeholder]

- Findings / key demographic for business to focus on

## Rank
- Overview
- Plot
- Findings / key areas for business to focus on / disparity between areas

## Map
- Overview
There are huge variations around the country, whether they be measured by
NHS Health Board or the more granular Local Authority areas.  Poverty is one
factor along with an urban/rural divide.  Essentially the areas with the lowest
proportion of smokers are more rural, and more affluent.  The opposite is also true.

- Findings / key areas for business to focus on
More work needs to be in less affluent areas, especially in the traditional
industrial areas,  to further reduce the rate of smoking.


# Key findings - Drug abuse | Jonathan

## The data
- Scottish Drug Misuse Database
- https://www.opendata.nhs.scot/
- Published by Public Health Scotland - the national public health body for Scotland
- Why was this data selected?


## Trend by gender  

```{r, message=FALSE, fig.width=9,fig.height=5}
sdmd_combined_plus_zones %>%
  filter(sex != "All") %>%
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>%
  filter(year != "2013/14") %>%
  arrange(year) %>%
  group_by(year, sex) %>%
  summarise(number_assessed_total = sum(number_assessed)) %>%
  ggplot(aes(x = year, y = number_assessed_total, group = sex, colour = sex)) +
  geom_point() +
  geom_line() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Numbers assessed at specialist drug treatment centres over time",
    subtitle = "By gender\n",
    x = "\nNumber assessed\n",
    y = "\nYear\n",
  ) +
  color_theme()
```

## Trend by age group  

```{r, fig.width=9,fig.height=5}
sdmd_combined_plus_zones %>%
  filter(age != "All" &
         age != "Unknown" &
         year != "2013/14") %>%
  filter(number_assessed != 0) %>%
  filter(!is.na(number_assessed)) %>%
  arrange(year) %>%
  group_by(year, age) %>%
  summarise(number_assessed_total = sum(number_assessed), .groups = "drop") %>%
  ggplot(aes(x = year, y = number_assessed_total, group = age, colour = age)) +
  geom_point() +
  geom_line() +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1)) +
  labs(
    title = "Numbers assessed at specialist drug treatment centres over time",
    subtitle = "By age group\n",
    x = "\nYear\n",
    y = "\nNumber of people assessed\n"
    ) +
  color_theme()
```


## Rank  

```{r, message=FALSE, fig.width=9,fig.height=5}
sdmd_combined_plus_zones %>% 
  filter(type == "NHS Health Board",
             year == "2018/19",
             sex == "All",
             name != "Shetland" & 
             name != "Western Isles") %>%
      group_by(name) %>% 
      summarise(total_number_assessed = sum(number_assessed)) %>% 
      arrange(desc(total_number_assessed)) %>% 
      ggplot() +
      aes(x = reorder(name, total_number_assessed), y = total_number_assessed) +
      geom_col() +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
      labs(
        title = "Numbers assessed by area",
        subtitle = "NHS Health Boards\n",
        x = "Data Zone",
        y = "\nNumber 
        assessed\n"
      )+
  color_theme()
```



## Conclusion | Jonathan
- Recap brief
- Recap of the main focus points for each of the three areas

# Dashboard Demonstration | Prathiba & Paddy


# Questions
